\section{Time and Projection}

\subsection{Problem Class}

Time is commonly treated as a fundamental parameter.
Within the MACHWERK framework, time is examined as a
\emph{derived, projected quantity} rather than a primitive entity.

The question addressed here is not how time evolves,
but under which formal conditions temporal statements
are physically admissible.

\subsection{Relational Formulation}

Let $\mathcal{U}$ denote the relational full space.
No element of $\mathcal{U}$ is assumed to possess an intrinsic time coordinate.

Temporal ordering, duration, and simultaneity are not defined in $\mathcal{U}$.

All temporal structure must arise through projection.

\subsection{Projection into an Observable Domain}

Let
\[
\Pi_t : \mathcal{U} \rightarrow m_2
\]
be a projection into an observable calculation domain $m_2$,
where time-like parameters are defined operationally.

The projection assigns to a relational configuration
a time parameter based on process comparison
(e.g. clock rates, decay rates, oscillatory references).

\subsection{Injectivity Analysis}

Injectivity of $\Pi_t$ requires that
\[
\forall u_1 \neq u_2 \in \mathcal{U} :
\Pi_t(u_1) \neq \Pi_t(u_2)
\]

In practice, multiple distinct relational configurations
may yield identical temporal readings.

This occurs whenever different underlying process structures
produce indistinguishable clock behavior.

Therefore, $\Pi_t$ is generically non-injective.

\subsection{Black Boundary Identification}

The set
\[
\Sigma_t := \{ u \in \mathcal{U} \mid \Pi_t \text{ is non-injective} \}
\]
defines the black boundary associated with temporal description.

Within $\Sigma_t$, temporal values remain computable,
but their interpretation as unique relational states is lost.

\subsection{Admissibility via CRA}

A temporal statement $S(t)$ is physically admissible if and only if
it depends solely on the projected observable $t \in m_2$
and is invariant under changes within the non-injective fiber of $\Pi_t$.

Statements that distinguish between relational states
inside $\Sigma_t$ violate the CRA axiom
and are physically inadmissible.

\subsection{Formal Conclusion}

Time is not a fundamental coordinate in MACHWERK.

It is a projection-dependent, domain-limited parameter whose
physical meaning is guaranteed only within injective regions.

Beyond the black boundary, time remains a valid computational label,
but ceases to support unique physical interpretation.
